# Machine Learning Project - Human Activity Recognition
Silviu Golumbeanu
August 22nd, 2015

## Overview 
Using the data available at  http://groupware.les.inf.puc-rio.br/har for the Weight Lifting Execrcise we are predicting the manner in which they did the exercise. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways and the data was captured from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

## Getting and Cleaning Data
### Loading Data
Loading the training and testing datasets:
```{r}
training_url <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testing_url <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training_raw <- read.csv(url(training_url), na.strings=c("NA","#DIV/0!",""))
testing_raw <- read.csv(url(testing_url), na.strings=c("NA","#DIV/0!",""))
dim(training_raw)
```
Loading libraries:
```{r}
library(caret)
library(rpart)
library(randomForest)
```
### Cleaning the Data
Determing the variables that would be the best predictore. We are selecting the variables for belt, arm, dumbbell, and forearm that do not have all NA values.
```{r}
NAs <- sapply(training_raw, function (x) any(is.na(x) | x == ""))
is_predictor <- !NAs & grepl("belt|[^(fore)]arm|dumbbell|forearm", names(NAs))
predictors <- names(NAs)[is_predictor]
vars <- c("classe", predictors)
training <- training_raw[, vars]
dim(training)
testing <- testing_raw[, predictors]
dim(testing)
```
Making the "classe" variable a factor:
```{r}
training$classe <- as.factor(training$classe) 
```
### Partioning the Training Set
We are splitting the Training data set into two data sets, 60% for training and 40% for testing:
```{r}
set.seed(12345)
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]
myTesting <- training[-inTrain, ]
```

## Model Building
### Using ML algorithms and decision trees for prediction
```{r}
modFit1 <- rpart(classe ~ ., data=myTraining, method="class")
```
Prediction:
```{r}
predictions1 <- predict(modFit1, myTesting, type = "class")
```
Sample Error and Cross Validation:
```{r}
confusionMatrix(predictions1, myTesting$classe)
```
### Using Random Forests
```{r}
modFit2 <- randomForest(classe ~. , data=myTraining)
```
Prediction:
```{r}
predictions2 <- predict(modFit2, myTesting, type = "class")
```
Sample Error and Cross Validation:
```{r}
confusionMatrix(predictions2, myTesting$classe)
```
## Generate file for submission
Use the Random Forests model on the provided test set:
```{r}
predictions3 <- predict(modFit2, testing, type = "class")
```
Generate the file for submission:
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictions3)
```

B2)
```
edictionsB2)
```
